# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
setwd("~/Desktop/Reproducible_Research/PA1")
df <- read.csv("activity.csv")
```

Have a look at the first 5 lines of the data
```{r}
head(df)
```

## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day
```{r}
day_sum <- tapply(df$steps, df$date, sum, na.rm = TRUE)
hist(day_sum, breaks = 10, main = "The histogram of the total number of steps taken each day")
```

calculate the mean and median total number of steps taken per day
```{r}
mean(day_sum)
median(day_sum)
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
five_minute_mean <- tapply(df$step, df$interval, mean, na.rm = TRUE)
plot(df$interval[1:288], five_minute_mean, type = 'l', xlab = "5 minutes interval in one day", ylab = "average number of steps taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max <- max(five_minute_mean)
index <- match(max, five_minute_mean)
max_point <- five_minute_mean[index]
max_interval <- as.numeric(names(max_point))
max_interval
```
So, the `r max_interval`th 5-minute interval contains the maxinum number of steps.
## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
missing_values <- is.na(df[,1])
number_NA <- sum(missing_values)
number_NA
```
So, the total missing value number is `r number_NA`

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I choose to fill the NAs with the mean for that 5-minute interval:
```{r}
five_minute_mean <- tapply(df$step, df$interval, mean, na.rm = TRUE)
df[missing_values, 1] <- five_minute_mean[df[missing_values, 3] %/% 100 * 12 + (df[missing_values, 3]%%100) /5 + 1]

head(five_minute_mean)
head(df)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
df2 <- df
missing_values2 <- is.na(df2[,1])
number_NA2 <- sum(missing_values2)
number_NA2

```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.  
```{r}
day_sum2 <- tapply(df2$steps, df2$date, sum)
hist(day_sum2, breaks = 10, main = "The histogram of the total number of steps taken each day (NA filled)")

mean(day_sum2)
median(day_sum2)
```
Q: Do these values differ from the estimates from the first part of the assignment?  
A: So the histogram, mean and median do differ from the first part when NAs were not filled.  
Q: What is the impact of imputing missing data on the estimates of the total daily number of steps?  
A: Making the data less biased.   



## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r cache = TRUE}
df2$day <- weekdays(as.Date(df2$date))
head(df2)

for (i in 1:length(df2$day)) {
        if (df2$day[i] == "Monday" | df2$day[i] == "Tuesday" | df2$day[i] == "Wednesday" | df2$day[i] == "Thursday" | df2$day[i] == "Friday") {
                df2$day[i] = "Weekday"
        }
        else {
                df2$day[i] = "Weekend"
        }
        }

head(df2)
df2_weekday <- df2[df2$day == "Weekday",]
df2_weekend <- df2[df2$day == "Weekend",]

```


Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
five_minute_mean_weekday <- tapply(df2_weekday$step, df2_weekday$interval, mean)
five_minute_mean_weekend <- tapply(df2_weekend$step, df2_weekend$interval, mean)
par(mfrow = c(2, 1))
plot(df$interval[1:288], five_minute_mean_weekday, type = 'l', xlab="", ylab = "number of steps", main = "weekdays")
plot(df$interval[1:288], five_minute_mean_weekend, type = 'l', xlab="interval", ylab = "", main = "weekends")
```

